---
title: "Swisstransplant cookbook for R"
author:
- name: Simon Schwab
  email: simon.schwab@swisstransplant.org
date: "Last updated: `r format(Sys.Date())`"
output: 
  html_document:
    code_folding: hide
    df_print: paged
    toc: true
    toc_depth: 2
    toc_float: true
subtitle: Research reporting
---

<style>
.list-group-item.active, .list-group-item.active:focus, .list-group-item.active:hover {
background-color: #e5005c;
}

a {
color: #e5005c;
}

h1.title, h1, h2 {
color: #2a548a;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Swisstransplant style graphics

This cookbook demonstrates the usage of the Swisstransplant (SWT) R package `swt`
to make high-quality publication-ready graphics in our in-house style.

## Load libraries
```{r, message=FALSE}
library(ggplot2)
library(gridExtra)
library(reshape2)
library(swt)
library(knitr)
```

## SWT colors
```{r fig.height=2, fig.width=8}
swtcol = swt_colors()

barplot(rep(1,10), axes=FALSE, col=c(swtcol$blue.swt,
                                     swtcol$turkis.tpx,
                                     swtcol$yellow.donation, 
                                     swtcol$strongred.akzent,
                                     swtcol$red.liver,
                                     swtcol$darkyellow.kidney,
                                     swtcol$green.pancreas,
                                     swtcol$lightblue.lungs,
                                     swtcol$beige.intestine,
                                     swtcol$pink.heart
),
names.arg = c("blue\nSWT", "turkis\nTpx", "yellow\ndonat", "strongr\nakzent", 
              "red\nliver", "darkylw\nkidney", "green\npancr", "lightb\nlungs",
              "beige\nintest", "pink\nheart")
)
```

The `swt_color` object provides an user-friendly access to the SWT
color scheme,  e.g. `swtcol$blue.swt` or `swtcol$sky.lungs`.


## Basic plots

```{r fig.height=5, fig.width=5}
set.seed(1980)
n = 100
var1 = c(rnorm(n/2, mean=0), rnorm(n/2, mean=1) )
d = data.frame(var1 = var1,
               var2 = var1 + rnorm(n, sd = 0.4),
               group = as.factor(rep(c("abc", "mno" ), each=n/2))
)

p1 = ggplot(d, aes(x=group, y=var1, group=group, col=group)) + 
  geom_boxplot(fill=swtcol$grey.bg) + 
  geom_point(size=2, position = position_jitter(height = 0, width = 0.15),
             col=swtcol$beige.intestine) +
  scale_color_manual(values = c(swtcol$blue.swt,
                                swtcol$pink.heart)) +
  swt_style(legend_position = "none", grey_theme = TRUE)

p2 = ggplot(d, aes(x=group, y=var1, group=group, col=group)) + 
  geom_boxplot() + 
  geom_point(size=2, shape=1, position = position_jitter(height = 0, width = 0.15),
             col=swtcol$beige.intestine) +
  scale_color_manual(values = c(swtcol$turkis.tpx,
                                swtcol$darkyellow.kidney)) +
  swt_style(legend_position = "none")

p3 = ggplot(d, aes(x=var2, fill=group)) + 
  geom_histogram(bins=10) +
  scale_fill_manual(values = c(swtcol$lightblue.lungs,
                               swtcol$beige.intestine)) +
  swt_style(grey_theme = FALSE)

p4 = ggplot(d, aes(x=var2, y=var1, group=group, col=group)) + 
  geom_point(size=2) +
  scale_color_manual(values = c(swtcol$blue.swt,
                                swtcol$yellow.donation)) +
  swt_style(legend_position = "top", grey_theme = TRUE)

grid.arrange(p1, p2, p3, p4, nrow = 2, ncol = 2)
```


## Make a line Plot

### Example data from annual report
```{r}
# Data taken from Annual Report 2020 (p. 32)
table3.2 = t(array(c(
  13.3,17.2,18.6,18.4,17.0,
  11.5,12.6,14.9,11.7,11.2,
  1.8,4.6,3.8,6.7,5.8), dim = c(5,3)))

colnames(table3.2) = 2016:2020
rownames(table3.2) = c("Total", "DBD", "DCD")
kable(table3.2)
```

Create a `data.frame` for the plot.
```{r}
data3.2 = melt(table3.2, varnames = c("Gruppe", "Jahr"), value.name = "Anzahl")
```

### Plot using SWT style
```{r fig.height=4, fig.width=5}
number_height = -0.8

ggplot(data3.2, aes(x=Jahr, y=Anzahl, col=Gruppe, group=Gruppe)) +
  
  # plot line with numbers
  geom_line(data = data3.2, size=1) +
  geom_text(data = data3.2, aes(label=Anzahl), vjust=number_height,
            col="black", size=5) +
  
  # some adjustments (colors, axes, etc)
  scale_color_manual(values=c(swtcol$strongred.akzent,
                              swtcol$yellow.donation,
                              swtcol$blue.swt)) +
  scale_y_continuous(breaks = seq(0,22,4), limits = c(0,22)) +
  
  ylab("pmp") +
  labs(title="Title",
       subtitle = "Subtitle"
  ) +
  swt_style(grey_theme = TRUE) +
  theme(legend.position = "top")
``` 

## Make a barplot with lines

### Example data from annual report
```{r}
# Data taken from Annual Report 2020 (p. 31)
table3.1 = t(array(c(
  96,106,126,100,96,
  15,39,32,57,50), dim = c(5,2)))

table3.1.totals = array(colSums(table3.1), dim = c(1,5))
colnames(table3.1) = 2016:2020
colnames(table3.1.totals) = 2016:2020
rownames(table3.1) = c("DBD", "DCD")
rownames(table3.1.totals) = c("Total")
kable(table3.1)
```

Create data.frame for the plot. Because the totals are in a separate line plot, we
create a separate `data frame` accordingly.
```{r}
data3.1 = melt(table3.1, varnames = c("Gruppe", "Jahr"), value.name = "Anzahl")
data3.1.totals = melt(table3.1.totals, varnames = c("Gruppe", "Jahr"),
                      value.name = "Anzahl")
```

```{r fig.height=4, fig.width=5}
number_height = -0.8
bar_with = 0.5

ggplot(data3.1, aes(x=Jahr, y=Anzahl, fill=Gruppe, group=Gruppe)) +
  
  # plot bars with numbers
  geom_bar(data = data3.1, stat="identity", position="dodge", width=bar_with) +
  geom_text(data = data3.1, aes(label=Anzahl), vjust=number_height,
            position = position_dodge(width=bar_with)) +
  
  # plot line with numbers
  geom_line(data = data3.1.totals, col = swtcol$strongred.akzent, size=1) +
  geom_text(data = data3.1.totals, aes(label=Anzahl), vjust=number_height,
            position = position_dodge(width=bar_with), col=swtcol$strongred.akzent) +
  
  # some adjustments (colors, axes, etc)
  scale_fill_manual(values=c(swtcol$yellow.donation, 
                             swtcol$blue.swt,
                             swtcol$strongred.akzent)) +
  scale_y_continuous(breaks = seq(0,180,20), limits = c(0,180)) +
  ylab("Personen") +
  swt_style(grey_theme = TRUE) 

``` 
