---
title: "Swisstransplant cookbook for R"
author:
- name: Simon Schwab
  email: simon.schwab@swisstransplant.org
output: 
  html_document:
    code_folding: hide
    df_print: paged
    toc: true
    toc_depth: 3
    toc_float: true
    css: "swt.css"
subtitle: Statistical reporting in our in-house style
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
This cookbook demonstrates the usage of the Swisstransplant R package `swt`
to make high-quality reports and publication-ready graphics in our in-house style.

```{r, message=FALSE}
library(ggplot2)
library(cowplot)
library(reshape2)
library(knitr)
library(swt)
```

# Header 1
This is a level-1 header.

## Header 2
This is a level-2 header.

### Header 3
This is a level-3 header.

# Colors

The `swt_color` object provides an user-friendly access to the SWT
color scheme.

```{r fig.height=1.2, fig.width=8}
col = swt_colors()
par(mfrow=c(1,1), mai=c(0.5,0.1,0.2,0.1))
barplot(rep(1,11), axes=FALSE, col=c(col$blue.dark,
                                     col$blue.swt,
                                     col$turkis.tpx,
                                     col$yellow.donation, 
                                     col$strongred.akzent,
                                     col$red.liver,
                                     col$darkyellow.kidney,
                                     col$green.pancreas,
                                     col$lightblue.lungs,
                                     col$beige.intestine,
                                     col$pink.heart
),
names.arg = c("blue\ndark", "blue\nSWT", "turkis\nTpx", "yellow\ndonat", "strongr\nakzent", 
              "red\nliver", "darkylw\nkidney", "green\npancr", "lightb\nlungs",
              "beige\nintest", "pink\nheart")
)
```

## Single hue palettes

The `swt_color` object also includes single hue palettes with three additonal color strengths, 75%, 50%, and 25%.

```{r fig.height=7, fig.width=5}
par(mfrow=c(10,1), mai=c(0.1,0.1,0.2,0.1))

barplot(rep(1,4), axes=FALSE, col=c(col$pal.blue.swt[1],
                                    col$pal.blue.swt[2],
                                    col$pal.blue.swt[3],
                                    col$pal.blue.swt[4]),
        names = c("100%", "75%", "50%", "25%")
)

barplot(rep(1,4), axes=FALSE, col=c(col$pal.turkis.tpx[1],
                                    col$pal.turkis.tpx[2],
                                    col$pal.turkis.tpx[3],
                                    col$pal.turkis.tpx[4])
)

barplot(rep(1,4), axes=FALSE, col=c(col$pal.yellow.donation[1],
                                    col$pal.yellow.donation[2],
                                    col$pal.yellow.donation[3],
                                    col$pal.yellow.donation[4])
)

barplot(rep(1,4), axes=FALSE, col=c(col$pal.strongred.akzent[1],
                                    col$pal.strongred.akzent[2],
                                    col$pal.strongred.akzent[3],
                                    col$pal.strongred.akzent[4])
)

barplot(rep(1,4), axes=FALSE, col=c(col$pal.red.liver[1],
                                    col$pal.red.liver[2],
                                    col$pal.red.liver[3],
                                    col$pal.red.liver[4])
)

barplot(rep(1,4), axes=FALSE, col=c(col$pal.darkyellow.kidney[1],
                                    col$pal.darkyellow.kidney[2],
                                    col$pal.darkyellow.kidney[3],
                                    col$pal.darkyellow.kidney[4])
)

barplot(rep(1,4), axes=FALSE, col=c(col$pal.green.pancreas[1],
                                    col$pal.green.pancreas[2],
                                    col$pal.green.pancreas[3],
                                    col$pal.green.pancreas[4])
)

barplot(rep(1,4), axes=FALSE, col=c(col$pal.lightblue.lungs[1],
                                    col$pal.lightblue.lungs[2],
                                    col$pal.lightblue.lungs[3],
                                    col$pal.lightblue.lungs[4])
)

barplot(rep(1,4), axes=FALSE, col=c(col$pal.beige.intestine[1],
                                    col$pal.beige.intestine[2],
                                    col$pal.beige.intestine[3],
                                    col$pal.beige.intestine[4])
)
```


# Plots
## Scientific plots

```{r fig.height=5, fig.width=5}
set.seed(1980)
n = 200
var1 = c(rnorm(n/2, mean=0), rnorm(n/2, mean=3) )
d = data.frame(var1 = var1,
               var2 = var1 + rnorm(n, sd = 0.4),
               group = as.factor(rep(c("abc", "mno" ), each=n/2))
)

p1 = ggplot(d, aes(x=group, y=var1, group=group, col=group)) + 
  geom_boxplot(fill=col$grey.bg) + 
  geom_point(size=2, position = position_jitter(height = 0, width = 0.25),
             col=c(rep(col$pal.blue.swt[3], n/2), rep(col$pal.red.liver[3], n/2))) +
  scale_color_manual(values = c(col$blue.swt,
                                col$red.liver)) +
  swt_style(legend_position = "none", grey_theme = TRUE)

p2 = ggplot(d, aes(x=group, y=var1, group=group, col=group)) + 
  geom_boxplot() + 
  geom_point(size=2, position = position_jitter(height = 0, width = 0.25),
             col=c(rep(col$pal.turkis.tpx[3], n/2), rep(col$pal.darkyellow.kidney[3], n/2))) +
  scale_color_manual(values = c(col$turkis.tpx,
                                col$darkyellow.kidney)) +
  swt_style(legend_position = "none")

p3 = ggplot(d, aes(x=var2, fill=group)) + 
  geom_histogram(position = "identity", bins=20, alpha=0.5) +
  scale_fill_manual(values = c(col$lightblue.lungs,
                               col$beige.intestine)) +
  swt_style(grey_theme = FALSE)

p4 = ggplot(d, aes(x=var2, y=var1, group=group, col=group)) + 
  geom_point(size=2, alpha=0.5) +
  scale_color_manual(values = c(col$blue.swt,
                                col$yellow.donation)) +
  labs(title = "Title") +
  swt_style(legend_position = "top", grey_theme = TRUE)

plot_grid(p1, p2, p3, p4, nrow = 2, ncol = 2, labels = c("A", "B", "C", "D"))
```


## Line plot

```{r}
# Data taken from Annual Report 2020 (p. 32)
table3.2 = t(array(c(
  13.3,17.2,18.6,18.4,17.0,
  11.5,12.6,14.9,11.7,11.2,
  1.8,4.6,3.8,6.7,5.8), dim = c(5,3)))

colnames(table3.2) = 2016:2020
rownames(table3.2) = c("Total", "DBD", "DCD")
```

First, we have to create a `data.frame` for the plot.

```{r}
data3.2 = melt(table3.2, varnames = c("Gruppe", "Jahr"), value.name = "Anzahl")
```

```{r fig.height=4, fig.width=5}
number_height = -0.8

ggplot(data3.2, aes(x=Jahr, y=Anzahl, col=Gruppe, group=Gruppe)) +
  
  # plot line with numbers
  geom_line(data = data3.2, linewidth=1) +
  geom_text(data = data3.2, aes(label=Anzahl), vjust=number_height,
            col="black", size=4) +
  
  # some adjustments (colors, axes, etc)
  scale_color_manual(values=c(col$strongred.akzent,
                              col$yellow.donation,
                              col$blue.swt)) +
  scale_y_continuous(breaks = seq(0,22,4), limits = c(0,22)) +
  
  ylab("pmp") +
  labs(title="Title",
       subtitle = "Subtitle"
  ) +
  swt_style(grey_theme = TRUE) +
  theme(legend.position = "top")
``` 

## Barplot with lines

```{r}
# Data taken from Annual Report 2020 (p. 31)
table3.1 = t(array(c(
  96,106,126,100,96,
  15,39,32,57,50), dim = c(5,2)))

table3.1.totals = array(colSums(table3.1), dim = c(1,5))
colnames(table3.1) = 2016:2020
colnames(table3.1.totals) = 2016:2020
rownames(table3.1) = c("DBD", "DCD")
rownames(table3.1.totals) = c("Total")
```

We first create a data.frame for the plot. Because the totals are in a separate line plot, we
create a separate `data frame` accordingly.

```{r}
data3.1 = melt(table3.1, varnames = c("Gruppe", "Jahr"), value.name = "Anzahl")
data3.1.totals = melt(table3.1.totals, varnames = c("Gruppe", "Jahr"),
                      value.name = "Anzahl")
```

```{r fig.height=4, fig.width=5}
number_height = -0.8
bar_with = 0.5

ggplot(data3.1, aes(x=Jahr, y=Anzahl, fill=Gruppe, group=Gruppe)) +
  
  # plot bars with numbers
  geom_bar(data = data3.1, stat="identity", position="dodge", width=bar_with) +
  geom_text(data = data3.1, aes(label=Anzahl), vjust=number_height,
            position = position_dodge(width=bar_with)) +
  
  # plot line with numbers
  geom_line(data = data3.1.totals, col = col$strongred.akzent, linewidth=1) +
  geom_text(data = data3.1.totals, aes(label=Anzahl), vjust=number_height,
            position = position_dodge(width=bar_with), col=col$strongred.akzent) +
  
  # some adjustments (colors, axes, etc)
  scale_fill_manual(values=c(col$yellow.donation, 
                             col$blue.swt,
                             col$strongred.akzent)) +
  scale_y_continuous(breaks = seq(0,180,20), limits = c(0,180)) +
  ylab("Personen") +
  swt_style(grey_theme = TRUE) 
```

<div class="footer"> 
  <hr> 
  <p>Last updated: `r format(Sys.Date(), '%d. %B %Y')`</p>
</div>

```{r}
sessionInfo()
```


